---
title: '*jumbled* demonstrations'
output: pdf_document
author: Sasha D. Hafner
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Overview
This document demonstrates usage of some of the function in the jumbled repo, available from [github.com/sashahafner/jumbled](github.com/sashahafner/jumbled).

# Load functions

```{r}
ff <- list.files(pattern = '\\.R$')
for(i in ff) source(i)
```

# aggregate2
A wrapper for `aggregate` that accepts multiple functions and simpler arguments.
Does not accept formula notation.

Example from `aggregate` help file:
```{r}
aggregate(breaks ~ wool + tension, data = warpbreaks, mean)
```

To include sd and n, use `aggregate2`:
```{r}
aggregate2(warpbreaks, x = 'breaks', by = c('wool', 'tension'), 
           FUN = list(mean = mean, sd = sd, n = length))
```

Accepts multiple variables (as in `aggregate`).

```{r}
aggregate2(na.omit(airquality), x = c('Ozone', 'Temp'), by = 'Month', 
	   FUN = list(mean = mean, sd = sd, n = length))
```

# aggregate3

Similar, but uses formula notation.
Example from `aggregate` help file:

```{r}
aggregate(breaks ~ wool + tension, data = warpbreaks, mean)
```

To include sd and n, use `aggregate3`:

```{r}
aggregate3(warpbreaks, breaks ~ wool + tension, 
	   FUN = list(mean = mean, sd = sd, n = length))
```

For multiple response variables, use `cbind()`.

```{r}
aggregate3(airquality, cbind(Ozone, Temp) ~ Month, 
	   FUN = list(mean = mean, sd = sd, n = length))
```

So `Ozone + Temp ~ Month` doesn't work, because `aggregate()` can't handle it propertly.
It would be nice to address this limitation in the future.

# dfcombos
Something like `expand.grid` for data frames.
Can accept vectors too, but resulting name is poor.

```{r}
d1 <- data.frame(name = letters[1:5], x = 1.1)
d2 <- data.frame(b = 1:3)
dfcombos(d1, d2)

v1 <- c(TRUE, FALSE)
dfcombos(d1, d2, v1)
```

# dfsumm
Generate a data frame summary more detailed and compact than `summary` output.

```{r}
dfsumm(attenu)
```

Add date to check R v4.3 problem.

```{r}
attenu$date.time <- Sys.time()
```

```{r}
dfsumm(attenu)
```

Compare to `summary`.

```{r}
summary(attenu)
```

# interpm
Fill in missing observations for multiple columns via interpolation.
`interpm` calls `approx`.

```{r}
args(interpm)
```


```{r}
dat <- data.frame(time = 1:30, a = rnorm(30), b = rnorm(30), c = rnorm(30))
dat[5:10, -1] <- NA
dat[20:22, 'a'] <- NA

dat

dat2 <- interpm(dat, 'time', c('a', 'b', 'c'))

dat2

plot(a ~ time, data = dat)
points(a ~ time, data = dat2, cex = 0.5, col = 'blue')
```

Now woks for data.tables too.

```{r}
dat <- data.table::as.data.table(dat)
dat2 <- interpm(dat, 'time', c('a', 'b', 'c'))
```

```{r}
dat <- data.frame(time = rep(1:10, 3), group = rep(c('a', 'b', 'c'), each = 10), a = rnorm(30), b = rnorm(30), c = rnorm(30))
dat[5:9, -1:-2] <- NA
dat[c(20, 22), 'a'] <- NA

dat
```

```{r}
interpm(dat, 'time', c('a', 'b', 'c'), by = 'group')
```

```{r}
interpm(dat, 'time', c('a', 'b', 'c'), by = 'group', rule = 2)
```

```{r}
dat <- data.table::as.data.table(dat)
dat
```

```{r}
interpm(dat, 'time', c('a', 'b', 'c'), by = 'group')
```

```{r}
interpm(dat, 'time', c('a', 'b', 'c'), by = 'group', rule = 2)
```


# logaxis
Add log axis to base R plots.

# logistic
The logistic function for transformations.

# rbindf
Like `rbind` but data frame columns do not need to match.
From monitoR package.

# rounddf
Round complete data frames.

```{r}
dat <- data.frame(a = 1:10, b = rnorm(10), c = letters[1:10])
dat


rounddf(dat)
rounddf(dat, digits = c(0, 4))
rounddf(dat, digits = c(0, 4), func = signif)
rounddf(dat, digits = c(2, 2), func = signif)
```

Trailing zeroes are dropped when written out (although this does not show up in R console).
Avoid with `pad = TRUE`, which converts adds trailing zeroes and converts column to character.


```{r}
set.seed(124)
dat <- data.frame(a = 1:10, b = rnorm(10), c = letters[1:10])
dat
summary(dat)

rounddf(dat)
rounddf(dat, pad = TRUE)
dat <- rounddf(dat, pad = TRUE)
summary(dat)
```



## `ggsave2x`
Save a ggplot2 figure in more than one format in a single call.

```{r}
library(ggplot2)
ggplot(economics, aes(date, unemploy)) +
  geom_line(colour = "red")
```

```{r, eval=FALSE}
ggsave2x('economics', width = 5, height = 5)
```

Saves png and pdf by default, add more with `type` argument.
Use `...` optional arguments for more flexibility.


## `mintegrate`
Integrate *f*lux measurements for emission.

```{r}
source('mintegrate.R')
```

### 1. Linear

```{r}
x <- 0:10
y <- 0:10
plot(x, y)
```

Exact integral is `10 * 10 / 2 = 50`.

```{r}
mintegrate(x, y, 'midpoint')
mintegrate(x, y, 'left')
mintegrate(x, y, 'right')
mintegrate(x, y, 'trap')
```

Note differences on the way up.

```{r}
plot(0:10, x * y / 2, ylim = c(0, 60))
lines(0:10, mintegrate(x, y, 'midpoint'), col = 'orange')
lines(0:10, mintegrate(x, y, 'left'), col = 'red')
lines(0:10, mintegrate(x, y, 'right'), col = 'blue')
lines(0:10, mintegrate(x, y, 'trap'), col = 'green', lty = 2)
```

Leave out 0 (say first measurement is at time = 1).

```{r}
x <- 1:10
y <- 1:10
plot(x, y)
```

Exact integral depends on what occurred before t = 1.

```{r}
mintegrate(x, y, 'midpoint')
mintegrate(x, y, 'left')
mintegrate(x, y, 'right')
mintegrate(x, y, 'trap')
```

Can incorporate assumptions.

```{r}
mintegrate(x, y, 'midpoint', lwr = 0)
mintegrate(x, y, 'left', lwr = 0)
mintegrate(x, y, 'right', lwr = 0)
mintegrate(x, y, 'trap', lwr = 0, ylwr = 0)
```

### Nonlinear

```{r}
x <- 1:10
y <- exp(-x)
plot(x, y, type = 'l')
```

Exact integral from 1:10 is `exp(-10) - exp(-1)` = 0.3678.
From 0 it is 1.0.

```{r}
mintegrate(x, y, 'midpoint', value = 'total')
mintegrate(x, y, 'left', value = 'total')
mintegrate(x, y, 'right', value = 'total')
mintegrate(x, y, 'trap', value = 'total')
```

```{r}
plot(0:10, -exp(-(0:10)) + exp(-min(0:10)), col = 'red')
points(x, -exp(-x) + exp(-min(x)), ylim = c(0, 0.7))
lines(x, mintegrate(x, y, 'midpoint'), col = 'orange')
lines(x, mintegrate(x, y, 'left'), col = 'red')
lines(x, mintegrate(x, y, 'right'), col = 'blue')
lines(x, mintegrate(x, y, 'trap'), col = 'green', lty = 2)
```

None is perfect, but midpoint and trapezoid (identical in this implementation) are the best, only slightly overestimating.
Note that they all do poorly compared to a true integral that starts at 0 (red points).
This cannot really be helped--how could we infer the true high values of y close to 0 from these limited measurements?

```{r}
x <- 0:10
plot(x, exp(-x))
```

The `lwr` argument can extend the first rate back to 0 or any arbitrary starting point, which helps a bit.

```{r}
x <- 1:10
plot(0:10, -exp(-(0:10)) + exp(-min(0:10)), col = 'red')
points(x, -exp(-x) + exp(-min(x)), ylim = c(0, 0.7))
lines(x, mintegrate(x, y, 'midpoint', lwr = 0), col = 'orange')
```

But measurements are needed at or closer to 0 to do really well with this function.
Start at 0.

```{r}
x <- 0:10
y <- exp(-x)
plot(x, y, type = 'l')
```

```{r}
mintegrate(x, y, 'midpoint', value = 'total')
mintegrate(x, y, 'left', value = 'total')
mintegrate(x, y, 'right', value = 'total')
mintegrate(x, y, 'trap', value = 'total')
```

```{r}
plot(x, -exp(-x) + exp(-min(x)), ylim = c(0, 1.5))
lines(x, mintegrate(x, y, 'midpoint'), col = 'orange')
lines(x, mintegrate(x, y, 'left'), col = 'red')
lines(x, mintegrate(x, y, 'right'), col = 'blue')
lines(x, mintegrate(x, y, 'trap'), col = 'green', lty = 2)
```

We can prove that all methods become accurate with very high resolution.

```{r}
x <- 0:100 / 10
y <- exp(-x)
plot(x, y, type = 'l')
```

```{r}
mintegrate(x, y, 'midpoint', value = 'total')
mintegrate(x, y, 'left', value = 'total')
mintegrate(x, y, 'right', value = 'total')
mintegrate(x, y, 'trap', value = 'total')
```

```{r}
x <- 0:10000 / 1000
y <- exp(-x)
plot(x, y, type = 'l')
```

```{r}
mintegrate(x, y, 'midpoint', value = 'total')
mintegrate(x, y, 'left', value = 'total')
mintegrate(x, y, 'right', value = 'total')
mintegrate(x, y, 'trap', value = 'total')
```

```{r}
plot(x, -exp(-x) + exp(-min(x)), col = 'gray')
lines(x, mintegrate(x, y, 'midpoint'), col = 'orange')
lines(x, mintegrate(x, y, 'left'), col = 'red')
lines(x, mintegrate(x, y, 'right'), col = 'blue')
lines(x, mintegrate(x, y, 'trap'), col = 'green', lty = 2)
```

Note that data need not be sorted by x.

```{r}
x <- 0:10
y <- exp(-x)
```

```{r}
mintegrate(x, y, 'midpoint')
```

```{r}
x[1] <- 4
x[5] <- 0
y <- exp(-x)
```

```{r}
int <- mintegrate(x, y, 'midpoint')
plot(x, int)
```

## difftime


```{r}
now <- Sys.time()
x <- difftime(now, now - 1:10)
y <- exp(-as.numeric(x))
```

```{r}
int <- mintegrate(x, y)
plot(x, int)
```

```{r}
mintegrate(x, y, value = 'xy')
```

With different units, result will differ.
It is up to the user to make sure y and x have same time unit!

```{r}
x <- difftime(now, now - 1:10, units = 'hours')
y <- exp(-as.numeric(x * 3600))
```

```{r}
mintegrate(x, y, value = 'xy')
```

## Grouped

```{r}
x <- 0:10
y <- exp(-x)
```


```{r}
x <- c(x, x)
y <- c(y, y + 0.2)
g <- rep(c('a', 'b'), each = 11)
```

```{r}
int <- mintegrate(x, y, by = g)
plot(x, int, pch = g)
```


